---
title: "NINJA_reports_20250211"
author: "Casey Sakamoto"
date: "2025-02-10"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
    number_sections: true
---
# Objectives
- Settle on primary var for AKI Timing/Stage
- Confirm data quality: make sure death var is fixed
- Look into Supratherapeutic Ranges?

```{r dc, include = FALSE}
library(tidyverse)
library(knitr)
library(haven)
# upload data d
data_d <- read_csv("S:/Brinton/Nephrology/2024.03.06 Shih NINJA - RAI mephrotoxin exposures in ICU and clincal course retrospective cohort/data/data_d.csv")


##################################################################################################################################
# report of recordid, aki date, aki stage, highest stage, highest date, whether the dates/stages differ
# data_d_report = data_d %>% select(record_id, contains("aki"), contains("AKI"),time_flag_to_peak, time_flag_to_AKI, date_of_maximum_scr_during, start_date_for_ntmx_encoun)

# note there are 3 subj who had time_flag_to_peak missing, in code, john looked at date of max sccr_during and start date of ntmx encounter -- use these?
data_d_report = data_d %>% select(record_id, aki_episode_start_date,aki_episode_resolution_dat, initial_aki_stage,
                                  highest_aki_stage_for_this, time_flag_to_AKI,time_flag_to_peak,
                                  date_of_maximum_scr_during, start_date_for_ntmx_encoun)

# since initial aki is in form Stage 1 Stage 2 Stage 3 
# and highest is: 
# Stage 1 (0.3 increase in Daily Scr  or Daily Scr 1.5 - 1.9 times baseline),
# Stage 2 (Daily Scr 2 - 2.9 times baseline)
# Stage 3 (Daily Scr >= 3 times baseline or provision of dialysis for an AKI related reason) 

# we will just see if the string from initial is in the highest
data_d_report$FLAG_Stage <- !mapply(function(a, b) grepl(a, b, fixed = TRUE), data_d_report$initial_aki_stage, data_d_report$highest_aki_stage_for_this)

# there are 3 subj: 67, 79, 83 who are missing the time to highest and date of max scr; these subj will be flagged in addition to the subj where the dates do not match
data_d_report$FLAG_Dates = !ifelse(is.na(data_d_report$time_flag_to_peak), FALSE, data_d_report$time_flag_to_peak == data_d_report$time_flag_to_AKI)

# select vars
data_d_report = data_d_report %>% select(record_id, time_flag_to_AKI, time_flag_to_peak, initial_aki_stage, highest_aki_stage_for_this, FLAG_Stage, FLAG_Dates)

# Add labels to the variables
labels_vector <- c(
  "Record ID",
  "Days from Flag to AKI",
  "Days from Flag to Peak",
  "Initial AKI Stage",
  "Peak AKI Stage",
  "Different Initial/Peak AKI Stage",
  "Different Initial/Peak Dates"
)

names(labels_vector) <- names(data_d_report)
for (i in seq_along(data_d_report)) {
  attr(data_d_report[[i]], "label") <- labels_vector[i]
}

data_d_report <- labelled::set_variable_labels(
  data_d_report,
  record_id = "Record ID",
  time_flag_to_AKI = "Days from Flag to AKI",
  time_flag_to_peak = "Days from Flag to Peak",
  initial_aki_stage = "Initial AKI Stage",
  highest_aki_stage_for_this = "Peak AKI Stage",
  FLAG_Stage = "Different Initial/Peak AKI Stage",
  FLAG_Dates = "Different Initial/Peak Dates"
)



##################################################################################################################################
data_d_ranges = data_d %>% select(record_id, additional_comments, supratherapeutic_vancomyci, gentamicin_trough_2, tacrolimus_level_15, cyclosporine_level_300)
labels_vector2 = c("Record ID","Additional Comments", "Vancomycin > 15", "Gentamicin Trough > 2", "Tacromilus Level > 15", "Cyclosporine Level > 300")


names(labels_vector2) <- names(data_d_ranges)
for (i in seq_along(data_d_ranges)) {
  attr(data_d_ranges[[i]], "label") <- labels_vector2[i]
}

library(table1)
ranget1 = table1(~.,data=data_d_ranges %>% select(-record_id, -additional_comments))
##################################################################################################################################
```

# Report

## Subj with Different Stage/Dates initial vs Peak

The distribution for highest aki has fewer stage 1 AKI and more stage 2/3 AKI than the distribution for initial AKI
```{r, echo=FALSE}
# initial
kable(table(data_d_report$initial_aki_stage),col.names = c("",""))
# peak
kable(table(data_d_report$highest_aki_stage_for_this),col.names = c("",""))
```

- Of 72 Subjects, 11 (15%) had different stages for initial vs peak AKI, while 33* (46%*) had different dates for initial vs peak AKI.

  -- *3 Subjects (rid's:67,79,83) missing the date of highest SCR (rid 83 also missing a highest scr as well).

- For stage flags, all subj except 1 also had a flag for different dates 
  -- 10 subj total with both diff stage and dates 


```{r}
# report of aki differences
# total subj with different Stages
sum(data_d_report$FLAG_Stage)
# total subj with different dates
sum(data_d_report$FLAG_Dates)

# looking at stages flags
kable(data_d_report %>% filter(FLAG_Stage ==1),col.names = labels_vector)

# looking at dates flags
kable(data_d_report %>% filter(FLAG_Dates ==1),col.names = labels_vector)
```

## Death Variable fix confirm

At standing meeting it was mentioned in the Table1 the death variable factoring was likely switched. Confirming the rate is more in line of what we would expect now after recoding

- The proportion is now 13/72 = 18%

```{r}
# confirm death variable fixed
table(data_d$death)
```

## Supratherapeutic Ranges

At meeting it was mentioned that supratherapeutic ranges were of interest, and to investigate the different ranges within the "additional_comments" variable

- additional_comments doesn't look like ranges. Looking through the dataset I am only seeing a few dichotomous variables with what looks like supratherapeutic values. 
```{r, echo=FALSE}
# not what we are looking for with the supratherapeutic ranges?
kable(table(data_d$additional_comments),col.names = c("Additional Comments", "Freq"))
```

- below is a summary of the supratherapeutic variables i found in the dataset.

```{r, echo=FALSE}
# look at the supratherapeutic variables from the data dictionary
ranget1
```