---
title: "NINJA Report 2025-03-11"
author: "Casey Sakamoto"
date: "2025-03-11"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
    number_sections: false
---
# Objectives
- Examinine the distribution of days from flag to initial aki
- Examinine/Summarize additional comments for supratherapeutic treatment
  - Subcategorize ST Vanc comments into dose groups

```{r dc, include = FALSE}
library(tidyverse)
library(knitr)
library(haven)
library(table1)
# upload data d
data_d <- read_csv("S:/Brinton/Nephrology/2024.03.06 Shih NINJA - RAI mephrotoxin exposures in ICU and clincal course retrospective cohort/data/data_d.csv")

data_d$death = factor(data_d$death, labels = c("No", "Yes"))
##################################################################################################################################
# report of recordid, aki date, aki stage, highest stage, highest date, whether the dates/stages differ
# data_d_report = data_d %>% select(record_id, contains("aki"), contains("AKI"),time_flag_to_peak, time_flag_to_AKI, date_of_maximum_scr_during, start_date_for_ntmx_encoun)

# note there are 3 subj who had time_flag_to_peak missing, in code, john looked at date of max sccr_during and start date of ntmx encounter -- use these?
data_d_report = data_d %>% select(record_id, aki_episode_start_date,aki_episode_resolution_dat, initial_aki_stage,
                                  highest_aki_stage_for_this, time_flag_to_AKI,time_flag_to_peak,
                                  date_of_maximum_scr_during, start_date_for_ntmx_encoun)

# since initial aki is in form Stage 1 Stage 2 Stage 3 
# and highest is: 
# Stage 1 (0.3 increase in Daily Scr  or Daily Scr 1.5 - 1.9 times baseline),
# Stage 2 (Daily Scr 2 - 2.9 times baseline)
# Stage 3 (Daily Scr >= 3 times baseline or provision of dialysis for an AKI related reason) 

# we will just see if the string from initial is in the highest
data_d_report$FLAG_Stage <- !mapply(function(a, b) grepl(a, b, fixed = TRUE), data_d_report$initial_aki_stage, data_d_report$highest_aki_stage_for_this)

# there are 3 subj: 67, 79, 83 who are missing the time to highest and date of max scr; these subj will be flagged in addition to the subj where the dates do not match
data_d_report$FLAG_Dates = !ifelse(is.na(data_d_report$time_flag_to_peak), FALSE, data_d_report$time_flag_to_peak == data_d_report$time_flag_to_AKI)

# select vars
data_d_report = data_d_report %>% select(record_id, time_flag_to_AKI, time_flag_to_peak, initial_aki_stage, highest_aki_stage_for_this, FLAG_Stage, FLAG_Dates)

# Add labels to the variables
labels_vector <- c(
  "Record ID",
  "Days from Flag to AKI",
  "Days from Flag to Peak",
  "Initial AKI Stage",
  "Peak AKI Stage",
  "Different Initial/Peak AKI Stage",
  "Different Initial/Peak Dates"
)

names(labels_vector) <- names(data_d_report)
for (i in seq_along(data_d_report)) {
  attr(data_d_report[[i]], "label") <- labels_vector[i]
}

data_d_report <- labelled::set_variable_labels(
  data_d_report,
  record_id = "Record ID",
  time_flag_to_AKI = "Days from Flag to AKI",
  time_flag_to_peak = "Days from Flag to Peak",
  initial_aki_stage = "Initial AKI Stage",
  highest_aki_stage_for_this = "Peak AKI Stage",
  FLAG_Stage = "Different Initial/Peak AKI Stage",
  FLAG_Dates = "Different Initial/Peak Dates"
)



##################################################################################################################################
data_d_ranges = data_d %>% select(record_id, additional_comments, supratherapeutic_vancomyci, gentamicin_trough_2, tacrolimus_level_15, cyclosporine_level_300)
labels_vector2 = c("Record ID","Additional Comments", "Vancomycin > 15", "Gentamicin Trough > 2", "Tacromilus Level > 15", "Cyclosporine Level > 300")


names(labels_vector2) <- names(data_d_ranges)
for (i in seq_along(data_d_ranges)) {
  attr(data_d_ranges[[i]], "label") <- labels_vector2[i]
}

library(table1)
ranget1 = table1(~.,data=data_d_ranges %>% select(-record_id, -additional_comments))
##################################################################################################################################
```

```{r, include=FALSE}
#######################
# SUPRATHERAPEUTIC COMMENTS DATA
st_comments<- read.csv("S:/Brinton/Nephrology/2024.03.06 Shih NINJA - RAI mephrotoxin exposures in ICU and clincal course retrospective cohort/data/st_vanc_comment.csv")
# pull comments from supratherapeutic variables
varcom_list = c("supratherapeutic_vancomyci","cyclosporine_level_300","gentamicin_trough_2", "tacrolimus_level_15")

st_comments = st_comments %>% filter(Field %in% varcom_list) 
# table(st_comments$Comment, st_comments$Field)

# st_comments_v = st_comments %>% filter(Field =="supratherapeutic_vancomyci")
# st_comments_c = st_comments %>% filter(Field =="cyclosporine_level_300")
# st_comments_t = st_comments %>% filter(Field =="tacrolimus_level_15")


# Function to process comments
extract_values <- function(comment) {
  # Remove dose-related phrases (e.g., "4 doses", "3/31", "5/10")
  clean_comment <- str_replace_all(comment, "\\b\\d+\\s*doses?\\b", "") # Remove "X doses"
  clean_comment <- str_replace_all(clean_comment, "\\b\\d+/\\d+\\b", "") # Remove date-like "X/X"

  # Extract all relevant numerical values
  numbers <- str_extract_all(clean_comment, "\\b\\d+\\.?\\d*\\b") %>% unlist() %>% as.numeric()
  
  highest_value <- ifelse(length(numbers) > 0, max(numbers, na.rm = TRUE), NA)
  first_value <- ifelse(length(numbers) > 0, numbers[1], NA)
  multiple_values <- ifelse(length(numbers) > 1, 1, 0)
  
  # Missing flag is 1 if no valid numbers or if explicit missing indicators are present
  missing_flag <- ifelse(length(numbers) == 0 | str_detect(comment, "<|not detected|N/A|no levels available"), 1, 0)
  multiple_values <- ifelse(missing_flag == 1, NA, multiple_values)

  # Extract goal range (handles "goal was X-Y", "goal is X-Y", and "goal X-Y")
  goal_match <- str_match(clean_comment, "goal (?:was |is )?(\\d+)[-–](\\d+)")
  goal_min <- ifelse(!is.na(goal_match[2]), as.numeric(goal_match[2]), NA)
  goal_max <- ifelse(!is.na(goal_match[3]), as.numeric(goal_match[3]), NA)
  
  return(c(highest_value, first_value, multiple_values, missing_flag, goal_min, goal_max))
}


# Apply function to each comment
st_comments <- st_comments %>%
  rowwise() %>%
  mutate(
    highest_value = extract_values(Comment)[1],
    first_value = extract_values(Comment)[2],
    multiple_values = extract_values(Comment)[3],
    missing_flag = extract_values(Comment)[4],
    goal_min = extract_values(Comment)[5],
    goal_max = extract_values(Comment)[6]
  )

st_comments_vanc = st_comments %>% filter(Field == "supratherapeutic_vancomyci")%>%
  mutate(highest_group = case_when(
    highest_value >= 15 & highest_value < 20 ~ "15-<20",
    highest_value >= 20 & highest_value < 25 ~ "20-<25",
    highest_value >= 25 & highest_value < 30 ~ "25-<30",
    highest_value >= 30                     ~ "≥30",
    TRUE ~ NA_character_),
    
         first_group = case_when(
    first_value >= 15 & first_value < 20 ~ "15-<20",
    first_value >= 20 & first_value < 25 ~ "20-<25",
    first_value >= 25 & first_value < 30 ~ "25-<30",
    first_value >= 30                     ~ "≥30",
    TRUE ~ NA_character_))

###########################
```

```{r, echo = FALSE}
#hist(data_d_report$time_flag_to_peak)
#table(data_d_report$time_flag_to_peak)
#ggplot(data = data_d_report, aes(x = time_flag_to_AKI, y = time_flag_to_peak)) + geom_point() + theme_minimal() + geom_abline(slope = 1, intercept = 0)
```

# Distribution of Days from Flag to Initial AKI

The distribution of days from the initial flag to the onset of acute kidney injury (AKI) shows that most cases occur within Day 0 - Day 2, after which the occurrence sharply declines. Beyond Day 4, only a small number of cases are observed.

```{r, echo=F}
# overall distn of days from flag to AKI
hist(data_d_report$time_flag_to_AKI)
table(data_d_report$time_flag_to_AKI)
```

When broken down by Initial AKI stage, Stage 1 cases are the most common and tend to occur earlier, with the majority happening within the first three days. Stage 2 cases are also concentrated in this early period but are slightly more spread out. Stage 3 cases appear both in the early period and later, with some cases occurring as late as 14 days after flag. 

14 Stage 1's, 2 Stage 2's and 4 Stage 3 initial AKIs occured on the same day as the flag (Total N on Day 0 = 20). 

```{r, echo=F}
# by stage
# ggplot(data = data_d_report,
#        aes(x=initial_aki_stage,y = time_flag_to_AKI ,
#            group = initial_aki_stage, fill = initial_aki_stage)) +
#   geom_dotplot(binaxis = 'y', stackdir = "center", dotsize = 0.5, binwidth = 0.75) +
#   scale_y_continuous(breaks = seq(0,max(data_d_report$time_flag_to_AKI),by=1)) +
#   labs(x = "Initial AKI Stage", y = "Days from Flag to AKI")+ theme_minimal()


# Calculate counts for each group
group_counts <- data_d_report %>%
  group_by(initial_aki_stage) %>%
  summarise(n = n())

# Plot with group n's
ggplot(data = data_d_report,
       aes(x = initial_aki_stage, y = time_flag_to_AKI,
           group = initial_aki_stage, fill = initial_aki_stage)) +
  geom_dotplot(binaxis = 'y', stackdir = "center", dotsize = 0.5, binwidth = 0.75) +
  scale_y_continuous(breaks = seq(0, max(data_d_report$time_flag_to_AKI), by = 1)) +
  labs(x = "Initial AKI Stage", y = "Days from Flag to AKI") +
  theme_minimal() +
  geom_text(data = group_counts, aes(x = initial_aki_stage, y = max(data_d_report$time_flag_to_AKI) + 1, 
                                     label = paste0("n = ", n)), 
            vjust = 0, size = 5)


table(data_d_report$time_flag_to_AKI, data_d_report$initial_aki_stage)

```

# Summary of Supratherapeutic Vancomycin Levels

On average, the highest vancomycin level recorded among subjects was approximately 21.4, with values ranging from 15.5 to 37.8. A small proportion of cases had levels exceeding 30, while nearly half of the cases had highest levels between 20 and 25. The first recorded vancomycin levels tended to be slightly lower, averaging around 18.6, with a majority of cases falling between 15 and 20. 5 subjects had multiple values in a comment, while 3 comments had missing info and instead only had "<4 doses no trough".

```{r, echo=F}
table1(~highest_value + highest_group + first_value + first_group + factor(multiple_values) + factor(missing_flag), data = st_comments_vanc)
```

# Appendix

## Supratherapeutic Vancomycin Comments

```{r, echo=FALSE}
kable((st_comments_vanc %>% select(-Field, -User)))
```

## Non-Vanc Comments

Subjects 57 and 80 had both a ST Vanc comment and a non-vanc comment attached to ST flags. Subject 80 also had multiple non-vanc comments.

```{r, echo=FALSE}
kable((st_comments %>% filter(Field != "supratherapeutic_vancomyci")))
```



## AKI Days From Flag = 0

- list of subjects with 0 days from flag to AKI (n = 20)

```{r}
# list of subjects with 0 days from flag to AKI (n = 20)
data_d_report %>% filter(time_flag_to_AKI ==0) %>% select(record_id, time_flag_to_AKI, initial_aki_stage, FLAG_Stage, FLAG_Dates)
```

## AKI Distn Peak

```{r}
library(stringr)
data_d_report$highest_aki_stage_for_this = str_sub(data_d_report$highest_aki_stage_for_this, 1,7)
# Calculate counts for each group
group_counts <- data_d_report %>%
  group_by(highest_aki_stage_for_this) %>%
  summarise(n = n())

# Plot with group n's
ggplot(data = data_d_report,
       aes(x = highest_aki_stage_for_this, y = time_flag_to_peak,
           group = highest_aki_stage_for_this, fill = highest_aki_stage_for_this)) +
  geom_dotplot(binaxis = 'y', stackdir = "center", dotsize = 1, binwidth = 0.75) +
  scale_y_continuous(breaks = seq(0, max(data_d_report$time_flag_to_peak, na.rm=T), by = 1)) +
  labs(x = "Peak AKI Stage", y = "Days from Flag to Peak AKI") +
  theme_minimal() +
  geom_text(data = group_counts, aes(x = highest_aki_stage_for_this, y = max(data_d_report$time_flag_to_peak) + 1, 
                                     label = paste0("n = ", n)), 
            vjust = 0, size = 5)
# n = 15 day 0's
```