---
title: "NINJA_reports_20250211"
author: "Casey Sakamoto"
date: "2025-02-10"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
    number_sections: false
---
# Objectives
- Compare timing and stages of initial AKI and highest AKI
  - Determine which variables to use for timing and staging
- Verify death rate variable is fixed and new death rate makes sense
- Read in variables containing information about supratherapeutic measures and summarize ranges

```{r dc, include = FALSE}
library(tidyverse)
library(knitr)
library(haven)
# upload data d
data_d <- read_csv("S:/Brinton/Nephrology/2024.03.06 Shih NINJA - RAI mephrotoxin exposures in ICU and clincal course retrospective cohort/data/data_d.csv")

data_d$death = factor(data_d$death, labels = c("No", "Yes"))
##################################################################################################################################
# report of recordid, aki date, aki stage, highest stage, highest date, whether the dates/stages differ
# data_d_report = data_d %>% select(record_id, contains("aki"), contains("AKI"),time_flag_to_peak, time_flag_to_AKI, date_of_maximum_scr_during, start_date_for_ntmx_encoun)

# note there are 3 subj who had time_flag_to_peak missing, in code, john looked at date of max sccr_during and start date of ntmx encounter -- use these?
data_d_report = data_d %>% select(record_id, aki_episode_start_date,aki_episode_resolution_dat, initial_aki_stage,
                                  highest_aki_stage_for_this, time_flag_to_AKI,time_flag_to_peak,
                                  date_of_maximum_scr_during, start_date_for_ntmx_encoun)

# since initial aki is in form Stage 1 Stage 2 Stage 3 
# and highest is: 
# Stage 1 (0.3 increase in Daily Scr  or Daily Scr 1.5 - 1.9 times baseline),
# Stage 2 (Daily Scr 2 - 2.9 times baseline)
# Stage 3 (Daily Scr >= 3 times baseline or provision of dialysis for an AKI related reason) 

# we will just see if the string from initial is in the highest
data_d_report$FLAG_Stage <- !mapply(function(a, b) grepl(a, b, fixed = TRUE), data_d_report$initial_aki_stage, data_d_report$highest_aki_stage_for_this)

# there are 3 subj: 67, 79, 83 who are missing the time to highest and date of max scr; these subj will be flagged in addition to the subj where the dates do not match
data_d_report$FLAG_Dates = !ifelse(is.na(data_d_report$time_flag_to_peak), FALSE, data_d_report$time_flag_to_peak == data_d_report$time_flag_to_AKI)

# select vars
data_d_report = data_d_report %>% select(record_id, time_flag_to_AKI, time_flag_to_peak, initial_aki_stage, highest_aki_stage_for_this, FLAG_Stage, FLAG_Dates)

# Add labels to the variables
labels_vector <- c(
  "Record ID",
  "Days from Flag to AKI",
  "Days from Flag to Peak",
  "Initial AKI Stage",
  "Peak AKI Stage",
  "Different Initial/Peak AKI Stage",
  "Different Initial/Peak Dates"
)

names(labels_vector) <- names(data_d_report)
for (i in seq_along(data_d_report)) {
  attr(data_d_report[[i]], "label") <- labels_vector[i]
}

data_d_report <- labelled::set_variable_labels(
  data_d_report,
  record_id = "Record ID",
  time_flag_to_AKI = "Days from Flag to AKI",
  time_flag_to_peak = "Days from Flag to Peak",
  initial_aki_stage = "Initial AKI Stage",
  highest_aki_stage_for_this = "Peak AKI Stage",
  FLAG_Stage = "Different Initial/Peak AKI Stage",
  FLAG_Dates = "Different Initial/Peak Dates"
)



##################################################################################################################################
data_d_ranges = data_d %>% select(record_id, additional_comments, supratherapeutic_vancomyci, gentamicin_trough_2, tacrolimus_level_15, cyclosporine_level_300)
labels_vector2 = c("Record ID","Additional Comments", "Vancomycin > 15", "Gentamicin Trough > 2", "Tacromilus Level > 15", "Cyclosporine Level > 300")


names(labels_vector2) <- names(data_d_ranges)
for (i in seq_along(data_d_ranges)) {
  attr(data_d_ranges[[i]], "label") <- labels_vector2[i]
}

library(table1)
ranget1 = table1(~.,data=data_d_ranges %>% select(-record_id, -additional_comments))
##################################################################################################################################
```

# Report

## Subjects with Different Stage/Dates of Initial AKI vs Peak AKI

Of 72 total subjects, 33 subjects had a different date for Initial AKI and Highest AKI. 10 subjects had both a different date and a different stage, while 1 subject had only a different stage.

The Record IDs, Days from flag to the initial AKI, Days from flag to highest AKI, initial and highest AKI stages, and flags for different stages/peaks (True = different stage/peak) are displayed below for subjects with different stages/dates for their initial and highest AKI. 

- Table 1: Subjects With Stage Flags list the 11 individuals with different values for initial AKI and highest AKI. 

- Table 2: Subjects With Date Flags lists all individuals with day of AKI differences between initial AKI and highest AKI.
  - Note Record_ID 20 does not appear in this table as both stage and day differed for this individual.

```{r, echo=FALSE}
# looking at stages flags
kable(data_d_report %>% filter(FLAG_Stage ==1),col.names = labels_vector, caption = "T1: Subjects With Stage Flags")

# looking at dates flags
kable(data_d_report %>% filter(FLAG_Dates ==1),col.names = labels_vector, caption = "T2: Subjects With Date Flags")
```

### Question for Dr. Stenson: Which variable should we use to summarize AKI date and AKI stage, Initial or Highest AKI?

## Death Variable Fix Confirm

At standing meeting it was mentioned in the Table1 the death variable factoring was likely switched.
John changed the factoring in the analysis dataset, and the mortality rate is now 13/72 = 18%

### Question for Dr. Stenson: Does this new rate seem reasonable?

```{r, echo=FALSE}
# confirm death variable fixed
kable(table(data_d$death), col.names = c("Death Y/N", "Freq"))
```

## Supratherapeutic Ranges

In order to examine supratherapeutic drug levels, we were told to examine the "additional_comments" variable. This variable doesn't seem to contain supratherapeutic ranges, with a summary table listed below.

### Question for Dr. Stenson: Where is this information stored in the dataset?

```{r, echo=FALSE}
# not what we are looking for with the supratherapeutic ranges?
kable(table(data_d$additional_comments),col.names = c("Additional Comments", "Freq"))
```


```{r, include=FALSE}
# look at the supratherapeutic variables from the data dictionary
ranget1
```