---
title: "child_analyitic_cohort"
author: "Casey Sakamoto"
date: "2025-01-27"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(table1)
library(knitr)

# Set up paths and read data
# Adjust file paths to match your system
tab_1_person <- read.csv("S:/Sakamoto/Nephrology/Data/Table1 Person Table.csv")
tab_2_encount <- read.csv("S:/Sakamoto/Nephrology/Data/Table2 Encounter Table.csv")
tab_4_labs <- read.csv("S:/Sakamoto/Nephrology/Data/Table 4 Labs.csv")
tab_7_proc <- read.csv("S:/Sakamoto/Nephrology/Data/Table 7 Procedures Table.csv")
tab_8_adt <- read.csv("S:/Sakamoto/Nephrology/Data/Table 8 ADT Table.csv")

```

```{r dc, include=FALSE}
# Step 1: Initial data setup
AKI_COMPASS_DATA_T1 <- tab_1_person

# Step 2: Derive and clean up variables


# Create Diff_DOD and ID_DOD_89 variables
## THESE DONT EXIST ANYMORE IN CURRENT T1
# AKI_COMPASS_DATA_T1 <- AKI_COMPASS_DATA_T1 %>%
#   mutate(
#     Diff_DOD = abs(Days_from_DOB_to_Epic_DOD - Days_from_DOB_to_CDPHE_DOD),
#     ID_DOD_89 = ifelse(is.na(Days_from_DOB_to_Epic_DOD) & is.na(Days_from_DOB_to_CDPHE_DOD), NA,
#                   ifelse(Days_from_DOB_to_Epic_DOD == "" | Days_from_DOB_to_CDPHE_DOD == "", 1, 0))
#   )


# Step 3: Frequency and descriptive statistics
# Frequency table for missing demographics
freq_table_t1 <- AKI_COMPASS_DATA_T1 %>%
  summarise(across(c(Sex, race, ethnicity, Death_YN), ~sum(is.na(.)))) # none missing 

# Summary statistics for `DaysDOBtoEpicDeath`
summary_stats_t1 <- AKI_COMPASS_DATA_T1 %>%
  summarise(
    n = n(),
    nmiss = sum(is.na(DaysDOBtoEpicDeath)),
    mean = mean(DaysDOBtoEpicDeath, na.rm = TRUE),
    median = median(DaysDOBtoEpicDeath, na.rm = TRUE),
    q1 = quantile(DaysDOBtoEpicDeath, 0.25, na.rm = TRUE),
    q3 = quantile(DaysDOBtoEpicDeath, 0.75, na.rm = TRUE),
    min = min(DaysDOBtoEpicDeath, na.rm = TRUE),
    max = max(DaysDOBtoEpicDeath, na.rm = TRUE)
  )

# Step 4: Filter data for large differences
## THIS OBSOLETE WITH NO MORE CDPHE VAR TO COMPARE
# play_data_365 <- AKI_COMPASS_DATA_T1 %>%
#   filter(Diff_DOD > 365)
# 
# play_data_7 <- AKI_COMPASS_DATA_T1 %>%
#   filter(Diff_DOD > 7)

# Step 5: Subset data for analysis
#AKI_COMPASS_DATA_T2 <- tab_2_encount
AKI_COMPASS_DATA_T2 <- tab_2_encount %>%
  filter(ArbPersonID %in% AKI_COMPASS_DATA_T1$ArbPersonID)

# Step 6: Derive Age-based flags
first <- AKI_COMPASS_DATA_T2 %>%
  arrange(ArbPersonID, AgeDays) %>%
  group_by(ArbPersonID) %>%
  slice(1) %>%
  ungroup()

first_X_age <- first %>%
  mutate(
    FLAG_missing_age = is.na(AgeDays),
    FLAG_kid_age = ifelse(AgeDays <= 6570 & AgeDays > 0, TRUE, FALSE)
  )

# Step 7: Frequency and summary statistics for first age
age_flags_summary <- first_X_age %>%
  summarise(
    missing_age_flag = sum(FLAG_missing_age, na.rm = TRUE),
    kid_age_flag = sum(FLAG_kid_age, na.rm = TRUE)
  )

baseline_summary <- first_X_age %>%
  summarise(
    n = n(),
    nmiss = sum(is.na(BaselineResultValue)),
    mean = mean(BaselineResultValue, na.rm = TRUE),
    median = median(BaselineResultValue, na.rm = TRUE),
    q1 = quantile(BaselineResultValue, 0.25, na.rm = TRUE),
    q3 = quantile(BaselineResultValue, 0.75, na.rm = TRUE),
    min = min(BaselineResultValue, na.rm = TRUE),
    max = max(BaselineResultValue, na.rm = TRUE)
  )

# Step 8: Filter unreasonable BaselineResultValue
over_8 <- first_X_age %>%
  filter(BaselineResultValue > 8)


### COMBINE TABS 1 AND 2 using AKICOMPASS1 and FIRSTXAGE

analysis = full_join(AKI_COMPASS_DATA_T1, first_X_age, by = c("ArbPersonID"))

# add in unreasonable baseline result value flag
analysis = analysis %>% mutate(FLAG_Baselineresultvalue = ifelse(BaselineResultValue > 8 | is.na(BaselineResultValue), TRUE, FALSE))
# sum(analysis$FLAG_Baselineresultvalue) # 3443
# sum(is.na(analysis$BaselineResultValue)) # 3402 -- 41 values above 8

# filter out age 18 and age 0
analysis= analysis%>% filter(FLAG_kid_age == T)

# Step 9: Lab data analysis
AKI_COMPASS_DATA_T4 <- tab_4_labs %>%
  filter(ArbPersonID %in% AKI_COMPASS_DATA_T1$ArbPersonID)
AKI_COMPASS_DATA_T4 = left_join(AKI_COMPASS_DATA_T4,analysis, by=c("ArbPersonID"))

# not calcing well fix after hearing form john on creatine labels to use and how to stage
labs_cr <- AKI_COMPASS_DATA_T4 %>%
  filter(LabCompName %in% c("POCT Creatinine", "Creatinine, Serum/Plasma", "Creatinine Serum/Plasma - U")) %>%
  mutate(
    Stage1_flag = ifelse((DaysFromDOBtoCollection - AgeDays <= 2 & NumericValue - BaselineResultValue >= 0.3) |
                           (DaysFromDOBtoCollection - AgeDays <= 7 & NumericValue / BaselineResultValue >= 1.5), 1, 0),
    Stage2_flag = ifelse(NumericValue / BaselineResultValue >= 2, 1, 0),
    Stage3_flag = ifelse(NumericValue / BaselineResultValue >= 3 | NumericValue >= 4, 1, 0))
    
labs_cr = labs_cr %>% mutate(Stage = case_when(Stage1_flag == 1 ~ 1,
                                               Stage2_flag == 1 ~ 2,
                                               Stage3_flag == 1 ~ 3,
                                               TRUE ~ 0))


# Output frequency of stages
stage_summary <- labs_cr %>%
  group_by(Stage) %>%
  summarise(count = n())

#tables1,2 combined into person + first encounter in "analysis" dataset
####
# COMBINE ANALYSIS AND TABLE 4 HERE
####

# Step 10: Table 7, Nephrology consult

# may need to confirm date happens during admission
tab_7_consult = tab_7_proc %>% filter(ProcedureName == "CONSULT INPT NEPHROLOGY")
analysis = left_join(analysis, tab_7_consult)

# analysis %>% summarise(sum(!is.na(ProcedureName)))

# Step 11: Table 8: Admission, discharge dates etc
# lot of duplicate transfer in/out, lets just look at discharge/admission for now

# create admission and discharge col to easy compare
tab8_adm = tab_8_adt %>% select(ArbPersonID, ArbEncounterID, EventType, DaysFromDOBtoEventdate) %>% filter(EventType %in% c("Admission"))
tab8_dis = tab_8_adt %>% select(ArbPersonID, ArbEncounterID, EventType, DaysFromDOBtoEventdate) %>% filter(EventType %in% c("Discharge"))

tab8_adm_w <- tab8_adm %>% 
  pivot_wider(
    names_from = EventType, 
    values_from = DaysFromDOBtoEventdate,
    values_fn = min
  ) %>%
  mutate(across(where(is.character), ~ na_if(., "NULL"))) %>%
  mutate(across(where(is.character), as.numeric))

tab8_dis_w <- tab8_dis %>% 
  pivot_wider(
    names_from = EventType, 
    values_from = DaysFromDOBtoEventdate,
    values_fn = max
  ) %>%
  mutate(across(where(is.character), ~ na_if(., "NULL"))) %>%
  mutate(across(where(is.character), as.numeric))


# add in to analysis dataset
analysis = left_join(analysis, tab8_adm_w)
analysis = left_join(analysis, tab8_dis_w)

rm(tab8_adm, tab8_adm_w, tab8_dis, tab8_dis_w)
# number of missing admissions/discharge
# sum(is.na(analysis$Admission));sum(is.na(analysis$Discharge))
# create a flag for death at admission n = 16
analysis$FLAG_died_admission <- analysis$Admission == analysis$DaysDOBtoEpicDeath
# sum(analysis$FLAG_died_admission, na.rm = T) 

analysis$death_discharge_days_diff <- analysis$DaysDOBtoEpicDeath - analysis$Discharge

# Categorize the difference
analysis$mortality_after_discharge <- cut(analysis$death_discharge_days_diff,
                                          breaks = c(-Inf,0 , 30, 90, 365, 730, 1825, 3650, Inf),
                                          labels = c("Before Discharge","<30 days", "30-90 days", "90 days-1 year", "1-2 years", "2-5 years", "5-10 years", ">10 years"),
                                          right = FALSE)


```

```{r tab 2 7 workspace, include = FALSE}
# tab2 cleaning

## primary admission dx derivation

# examine what we are looking at
# 2944 different categories with 
Dx1 = as.data.frame(table(analysis$PrimaryDx))
# 523 have frequency of 5 or greater. 
# Dx1_g5 = Dx1 %>% filter(Freq >4)

# it looks like a lot of these categories can maybe be compressed using grepl or some sort of string matching function

## financial class

FC1 = as.data.frame(table(analysis$payorfinancialclass))
# this one seems more easy to clean up with a grepl string match -- confirm that this is ok with john

# tab 7 cleaning: 
# look for neph consult in procedures
consult_df = as.data.frame(table(tab_7_proc$ProcedureName))
# looks pretty clear cut -- only one category for neph consult -- double check with John that this makes sense
```

# Questions

## T1

- it seems like mortality after discharge/during admission can be derived, but would need the admission/discharge window dates -- see t8 question

## T2
- how to clean up t2 primary initial diagnosis PrimaryDx
- how to clean up financial class 

## T7 
- confirm consult is the only category in procedures table, right now we have 24 consults, 13 of which are in our analysis data

## T8
- Looks like some values are NULL even in the original dataset; missing 3303 admission dates and 2783 discharge dates. expected?



Vars: 
Label name	Var name	location

## DERIVED/READY TO PUT IN TABLE AS IS

Age at admission	
Sex (male)	
Race	Race	
Ethnicity	
In state (yes)	
Died (yes) 	

Died During admission	- derived as FLAG_died_admissions

Mortality after discharge	- now using epic DOD as only var we have now -- derived as 
  30 day		
  90 day		
  1 year		
  2 year		
  5 year		
  10 year
  
## WAITING ON EMAIL FROM DONIA (UPDATED DATA) AND GUIDANCE ON WHICH WEIGHT/HEIGHT TO USE (SEE TAB_3_CLEANING PROGRAM)

Weight on admission (kg)	?	Table 3 Flow Sheet 
Wt for age z score		NA - need age in months  Derived Table 3
Height on admission (cm)	?	Table 3 Flow Sheet
Height for age z score 		NA - need age in months  Derived Table 3
BMI (use CDC staging for categorical, also report as continuous		NA - need age in months  Derived Table 3
BSA (Body Surface Area)	?	Table 3 Flow Sheet
ECMO (yes)	?	Table 3? (Daniell emailed asking for this 4.7.2022)


## VARS TO CONFIRM QUESTIONS

Primary admission dx (will need to categorize) PrimaryAdmissionDiagnosis	Table 2 Encounters
Financial Class	FinancialClass	Table 2 Encounters
Nephrology consult	?	Table 7

## VARS TO FIND/DERIVE

Nephrotoxin	?	Table 5? - https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7393580/ - use this to identify exposure to neptox. exposures during hosp

Number		
  1		
  2		
  3		
  .>3	
  
Community acquired AKI, yes		? -- look at lab scr 180 days before admission, during admission and discharge
Time from admission to AKI in those with hospital acquired AKI (? Days or hours)		?
Prior history of KDIGO AKI 	?	?
Stage 1		
Stage 2		
Stage 3		




## NO LONGER EXIST IN CURRENT DATASET

Zip Code	Zip_code_3_digits	Table 2 Encounters
Urban vs suburban		Table 2 Encounters
Distance from CHCO	distance_from_hospital	Table 2 Encounters
  <20 miles		
  20-50 miles		
  50-80 miles		
  80-100		
, >100


# DESC TABLES

```{r t12}
#sum(analysis$FLAG_kid_age)
# 15070 out of 18904 had a T flag for kid age (between 1 and 6570 days)
# Variables of interest for T1

# ALL Vars t1 and t2
t1_all = table1(~ Sex+ race + ethnicity + DaysDOBtoEpicDeath + Death_YN + AgeDays +
         Hospital + State  + AKICohortYN + CRRTCohortYN + AKILabYN + AKIDxYN + BaselineResultValue, 
       data = analysis)

t1_all

table(analysis$FLAG_died_admission)
table(analysis$mortality_after_discharge)

```

SAS CODE Breakdown:
t1
- the four variables of interest for Table 1 are:
                Sex
                Race
                Ethnicity
                Deceased_yn
- cdphe vs epic dod discrepancy outdated due to no more cdphe dod variable exists
-- similarly no diffdod or dod89 vars
- injury description in appendix (very long)

t2
- only care about IDS in t2 that occur in t1 -- 19375 -> 18904 (matching n with tab1)
- take first age as age at admission, flag between 0 and 6570 (<18yo)
- look at those with baseline values; high baseline values = ~ 41, 3402 missing

t4
- 'POCT Creatinine','Creatinine, Serum/Plasma','Creatinine Serum/Plasma - U' taken from labcomp name(full list in appendix)
- below logic used to code flags using baseline value and lab value
  Stage1_flag = 0;
	Stage2_flag = 0;
	Stage3_flag = 0;
	day = DaysFromDOBtoCollection - AgeDays ;
	diff_scr = NumericValue - BaselineResultValue;
	ratio_scr = NumericValue/BaselineResultValue;

	IF Day <=2 Then DO;
	   if diff_scr >=0.3 THEN Stage1_flag = 1;
	END;
	IF Day <=7 THen Do;
		if ratio_scr >=1.5 THEN Stage1_flag = 1;
	END;

		if ratio_scr >=2 THEN Stage2_flag = 1;

		if ratio_scr >=3 THEN Stage3_flag = 1;
		if NumericValue>=4 THEN Stage3_flag = 1;

Stage = 0;
IF (Stage1_flag) THEN STAGE = 1;
IF (Stage2_flag) THEN STAGE = 2;
IF (Stage3_flag) THEN STAGE = 3;

- 


```{r appendix}
# labs
kable(table(AKI_COMPASS_DATA_T4$LabCompName))

# pay or financial class?
kable(table(analysis$payorfinancialclass))

# Primary admission description?
kable(table(analysis$PrimaryDx))
```

